### 2010-06-30 article.tex
First stuff getting put in

\documentclass[11pt]{article}
\usepackage[letterpaper]{geometry}
\usepackage{url}

\title{A Theoretical Approach to Virtual Memory}
\author{Peter Boothe\\
Manhattan College\\
\url{peter.boothe@manhattan.edu}}
\date{Submission to SODA 2011} % or maybe ALENEX2011

\newcommand{\heapsort}{{\sc HeapSort}}
\newcommand{\quicksort}{{\sc QuickSort}}
\newcommand{\bubblesort}{{\sc BubbleSort}}

\begin{document}
\maketitle

As memory requirements begin to outstrip available memory on a machine, the
operating system will allocate virtual memory to supplement the available
physical memory.  We present, argue for, and experimentally validate a simple
analysis technique which allows algorithms researchers to analyze the behavior
of an algorithm when using virtual memory.  In particular, the penalty for
fetching a page from virtual memory can be extreme, and can make
previously-small constants quite large. Thus, it would be of great use to be
able to easily estimate, for a given algorithm, the number of times the
algorithm will have to pause to retrieve virtual memory off the hard drive.

\section{The Method}

In the worst case, every memory access will have to go to the hard drive.
This, however, is not an informative worst-case analysis.  The worst-case page
fault predictions do not reflect the real-world behavior of virtual memory ---
they reflect a situation in which every cache access is a cache miss!  Instead,
our method is to simply count the number of non-adjacent memory and array
accesses.  Using this, we find that we also end up providing a quantitative
measure of ``cache friendliness'', which is a term often used, but seldomly
rigorously defined.  We will demonstrate our method using three classic sorting
techniques: \heapsort, \quicksort, and \bubblesort.

Our analysis of \heapsort is the easiest.  The \heapsort algorithm is simply 
\begin{verbatim}
heapsort(array, size)
    make-heap(array, size)
    for i = size-1 ... 0
        array[i] = remove-min(array, i+1)
    return array
\end{verbatim}

\end{document}